### 2019-10-16 tesis_hiram/Chapter1/chapter1.tex
final

Particle physics is a field of study of physics that studies the particles that create matter and its interactions to very small scale, even smaller than atoms. The particles analized in particle physics are the fundamental particles that compose the matter and the responsibles for all forces of nature. All thoses particles  and their properties are grouped in a theory called the Standard Model. Standard Model (SM) is a theory that describes the fundamental forces (except gravity) and classify the elementary particles in bosons (entire spin) and fermions (half spin).
The SM , created almost 50 years ago, has been the most successul theory in explain the phenomena in the subatomic world.
\\

One of the particles of the SM, the Higgs Boson , was recently discovered in 2012 in the ATLAS and CMS experiments at CERN. The discovery of Higgs boson in  proton proton collisions  marked a new era in the particle physics due to its characteristic  of giving the matter the property of mass, according to previous theories and predictions . Since then, people around the world are working in the Large Hadron Collider at CERN  in order to detect the Higgs bosons in higher energies and differents Higgs production processes. 
Also the impact of Higgs is also  important in cosmology,  where theoretical investigations relate the Higgs boson and the cosmic inflation.\\

In the recent years, there have been many studies in the Higgs production processes at CERN in pp collisions , such as gluon fusion (ggF), VBF, WH, ZH and $t\bar{t}$H and people have obtained consistent results in the search of Higgs in those processes. 
In this work, we will analize the production of Higgs boson in association with a
single top quark (tH) in proton-proton collisions with the CMS experiment of the LHC. 
The motivation of the work is this mechanism of production of the Higgs boson has not been observed before by any experiment. The exploration of Higgs production on the tH channel is subject relatively new. %The SM is used to study the tH channel and compare it with other Higgs production mechanisms results.
\\
Many channels are yet out of
reach experimentally. At the same time, many theoretical
proposals are yet to be put on test. It is possible that studies on the Higgs boson give evidence of small deviations from the SM  and test new physics that are beyond SM such as String Theory and  Supersymmetry. \\
%********************************** % Third Section  *************************************

@@

For the leptons we have electrons ($e$), muons ($\mu$), and taus ($\tau$) with charge -1 and their neutrinos $\nu_e, \nu_\mu,\nu_\tau$, are chargeless particles. Leptons have their respective antiparticle; same structure with opposite charge. 
 All of them have a spin (intrinsic angular momentum) of $\frac{1}{2}$\cite{griff}.% In case of neutrinos, so far experimentalists have found these have very small masses, but in the theoretical scheme, they are massless. \\
\\

 Quarks combine to form composite particles called hadrons. Quarks cannot be found in isolation, they always exist in form of hadrons that can be mesons or baryons. \\
Mesons , are bosons made up of a quark-antiquark pair, such as pion ($\pi$) and kaon ($K$). Mesons were theorized by Yukawa in 1937 and discovered in 1947 by Powell\cite{griff}.
Baryons are fermions made of 3 quarks. Examples of baryons are protons and neutrons. Both particles can interact via strong force.
\\

Quarks were first theorized by Murray Gell-Mann and 
George Zweig in 1964, were introduced as parts of an ordering scheme for hadrons, and there was little evidence for their physical existence until scattering experiments at the Stanford Linear Accelerator Center in 1968\cite{griff}.
There are 6 of them: up (1968), down (1968), strange (1968), charm (theorized in 1970, discovered 1974), bottom (theorized 1973, discovered 1977), top (theorized 1973, discovered 1995).
\\

Quarks have characteristics such as  electric charge, mass, color charge, and spin. The quarks up, charm and top have in common a charge of $\frac{2}{3}$, while the down , strange and bottom quarks have charge of $-\frac{1}{3}$. Among them , the top quark is the most massive of all, while the up and down are the least massive. 
The lepton and quarks have their respective antiparticle, which have the same mass, but inverted charge.\\  
Some of the particles in the SM are not stable, so they decays and form particles with lower masses. Other particles do not decay or their lifetimes are very long. The final states can be represented with a Feynman diagram according to the SM vertices\cite{mark}.

@@

The last group of the family are the bosons,which have a spin 1 and 0. 
There are 5 bosons: $W^{\pm}$, $Z$ with neutral charge are responsibles for weak force; photon, a massless particle, is responsible for electromagnetic force; gluon which also massless, is related to strong force and Higgs boson, recently discovered , is in charge of give the property of mass to particles via Higgs mechanism. Bosons let other particles to interact, even themselves, which is shown in the figure \ref{sm}. 
\\
The $W$, $Z$ and Higgs bosons, can interact with the most of the particles, while the gluon can only interact with quarks and itself.
Those bosons act as mediators in decays and creation of complex
particles, as mesons and baryons.\\

@@

\begin{table}[!htbp]
\caption{Table of particles in SM\cite{pd}. Stable means no decay, or lifetime almost infinite. Particles speeds at 0.998c}
\renewcommand{\arraystretch}{1.5}
\begin{tabular}{|c|c|c|c|c|p{2.5cm}|}
\hline 
Particle	& Mass (MeV/$c^2$)  &Charge  & spin &Lifetime (s)  & Distance in lifetime (meters) \\ 
	\hline 
Up ($u$)	& 2.2 & $\frac{2}{3}$ & $\frac{1}{2}$ & stable &  -\\ 
	\hline 
Charm ($c$)	& 1280 &$\frac{2}{3}$  &  $\frac{1}{2}$ & $ 1.1 \times 10^{-12}$ &  5.21$\times 10^{-3}$ \\ 
	\hline 
Top	($t$)& 173100 & $\frac{2}{3}$ & $\frac{1}{2}$  & 5$\times$10$^{-25}$ &2.37$\times 10^{-15}$   \\ 
	\hline 
Down ($d$)	& 4.6 &$-\frac{1}{3}$  & $\frac{1}{2}$  & Stable & - \\ 
	\hline 
Strange ($s$)	& 96 &$-\frac{1}{3}$  & $\frac{1}{2}$  &$1.24 \times 10^{-8}$  & 58.7 \\ 
	\hline

Bottom ($b$)	& 4180 &$-\frac{1}{3}$  &  $\frac{1}{2}$ &$1.3 \times 10^{-12}$   & 6.16 $\times 10^{-3}$\\ 
	\hline 
$W$ 	& 80379 &$\pm$1  & 1 & 3$\times$ 10$^{-25}$ &  1.42$\times 10^{-15}$\\ 
	\hline 
$Z$ & 91187.6 &0  & 1 & 3$\times$ 10$^{-25}$  &1.42$\times 10^{-15}$ \\ 
\hline
Photon ($\gamma$) & 0 &0  &  1&Stable  & - \\ 
\hline
Gluon ($g$)	& 0 &0  &  1&Stable  & - \\ 
	\hline 
Higgs ($H$)	& 125.18 &0  &  0& 1.56 $\times$ $10^{-22}$ & 7.39$\times 10^{-13}$ \\ 
	\hline 
Electron ($e$)& 0.511 & -1 &   $\frac{1}{2}$& Stable & - \\ 
	\hline 
Muon ($\mu$)	& 105.7 & -1  &  $\frac{1}{2}$ & 2.2$\times$10$^{-6}$ & 10419.85 \\ 
	\hline 
$\tau$	& 1776.86 &-1  &  $\frac{1}{2}$ & 2.9$\times$10$^{-13}$ & 1.37$\times 10^{-3}$\\ 
	\hline 
$\nu_e$	$\nu_\mu$ $\nu_\tau$& 0 & 0 &  $\frac{1}{2}$ & Stable &  -\\
	\hline 
\end{tabular} 
\label{SM table}
\end{table}
\newpage

\section{Higgs Boson}
The Higgs is a boson with spin 0 and chargeless. First theorized in 1964 by a group of theoretical physicist composed by Peter Higgs, Fran\c cois Englert, Robert Brout, Gerald Guralnik, C. Richard Hagen, and Tom Kibble in separated groups. They formulated the Higgs mechanism that explain the generation of mass for the particles . \\

@@

\section{Top quark}
The top quark  is the most massive particle  with a charge of 2/3 with spin of 1/2. First proposed by  M. Kobayashi and T. Maskawa in 1973 to explain observed CP violations in Kaon decay\cite{griff}. On 1995, in the Tevatron at Fermilab in Illinois, the top quark  was discovered via proton antiproton collisions with center of mass energy between 1.8 to 1.96 TeV in the CDF and D0 experiments.The top quark production process was via ggF, which generated a pair of $t\bar{t}$\cite{top} .
Since then,the properties of the top quark have been studied at the Tevatron Collider at Fermilab, and by the ATLAS and CMS experiments at the Large Hadron Collider (LHC) at CERN.\\
However, the top quark often receives special attention in new physics models because its mass requires near unity coupling to the Higgs boson\cite{top}.
As shown in table \ref{SM table}, the top quark lifetime is so small that they decay before they can hadronise, that is to say, the process to create hadrons from quarks and gluons.\\

@@

$W^i$ (i=1,2,3) are three  SU(2) gauge bosons and B is a U(1) gauge boson.\footnote{$\epsilon^{ijk}$ represents the Levi-Civita tensor, where $\epsilon^{ijk}= \begin{dcases}
	+1 \quad  \text{for even permutation} \\
	-1 \quad \text{for odd permutation}\\
	0 \quad \text{i=j, or j=k, or k=i } 
	\end{dcases}$.} In the end, those gauge bosons, after introduce the Higgs mechanism, become the $W^\pm$ , $Z$ and $\gamma$ vector boson
\cite{ew1}\cite{ew2}. %Z bosons  and photons will be formed by a combination of W and B fields.\\
\\

@@

we define the physical vector boson $W^-_\mu$. $W^+_\mu$ and $Z$ 
\begin{align} 
W^{\pm}_\mu=\frac{1}{\sqrt{2}} (W^1_\mu \mp iW^2_\mu) \label{smo}\\ 
Z_\mu=\frac{1}{\sqrt{g^2+g'^2}}\left(gW^3_\mu -g'B_\mu   \right) \label{zboson}
\end{align}

@@

The last part of the lagrangian is the Yukawa lagrangian
\begin{align}
\mathcal{L}_\text{yukawa}=\sum_{m,n}^{3}  \Gamma^u_{mn}\bar{q}_{m\text{,}L} \tilde{\Phi} u_{n\text{,}R}+\Gamma^d_{mn}\bar{q}_{m\text{,}L} \Phi d_{n\text{,}R}+\Gamma^e_{mn}\bar{l}_{m\text{,}L} \Phi e_{n\text{,}R}+h.c
\end{align}
The matrices $\Gamma_{mn}$ describe the so called Yukawa couplings between Higgs doublet $\Phi$ and the fermions.\footnote{Here $
	\tilde{\Phi}=i\sigma^2 \Phi^\dagger =\left(\begin{array}{c}
	\phi^{0^\dagger} \\
	-\phi^-
	\end{array} \right)
	% \Phi =\sqrt{-\frac{\mu^2}{2\lambda}}=\frac{v}{\sqrt{2}}  \qquad \lambda >0
	$ is necessary for the correct transformation in the case of up quarks. \\
h.c refers to hermitian conjugate terms.  }  The indices  m and n mean sum over the families. \\
%It is required to have two representations of Higgs fields
%with $Y=\frac{1}{2}$  and $\frac{1}{2}$ to give masses to  quarks and electrons.

@@

In tH process, Higgs boson is radiated from a single top quark as shown in \ref{psu}. tH process is a new field of study, given that there are few experiments and its production rate (probability) is small, so the detection of Higgs bosons in this channel opens to new discoveries.
\\

@@

In the table \ref{crt} shows the Higgs production cross section for pp collision and the number of events produced using a integrated luminosity of $35.9 fb^{-1}$ , that is the luminosity measured for the Run 2 from LHC on 2016\cite{pd}.

@@

Here we can see ggF process have the biggest cross section of all Higgs mechanism and generates the greatest number of events possibles. For the processes tH and $t\bar{t}$H, they have the smallest cross section, so the probability for the processes is low and generates a small number of events. 
\pagebreak
\footnotetext{Taken from \url{ https://twiki.cern.ch/twiki/bin/view/LHCPhysics/LHCHXSWGCrossSectionsFigures}}
\section{Higgs decay rates}
In particle physics,two important properties are the lifetime and the decay rates of a particle. The lifetime is related to the total decay rate $\Gamma$, that is the probability per unit of time a particle decay, 
\begin{align}
    dN=-\Gamma N dt
\end{align}
with N the original number of particles before the decay. From here is evident that number of particles left in a decay is

\begin{align}
N(t)=N_0 e^{-\Gamma t}
\end{align}

>>>>

Particle physics is a field of study of physics that studies the particles that create matter and its interactions to very small scale, even smaller than atoms. The particles analyzed in particle physics are the fundamental particles that compose the matter and responsible for all forces of nature. All those particles  and their properties are grouped in a theory called the Standard Model. Standard Model (SM) is a theory that describes the fundamental forces (except gravity) and classify the elementary particles in bosons (entire spin) and fermions (half spin).
The SM, created almost 50 years ago, has been the most successful theory in explain the phenomena in the subatomic world.
\\

One of the particles of the SM, the Higgs Boson, was recently discovered in 2012 in the ATLAS and CMS experiments at CERN. The discovery of Higgs boson in  proton proton collisions marked a new era in the particle physics due to its characteristic of giving the matter the property of mass, according to previous theories and predictions. Since then, people around the world are working in the Large Hadron Collider at CERN in order to detect the Higgs bosons in higher energies and different Higgs production processes. 
Also the impact of Higgs is also important in cosmology,  where theoretical investigations relate the Higgs boson and the cosmic inflation.\\

In the recent years, there have been many studies in the Higgs production processes at CERN in pp collisions, such as gluon fusion ($gg$F), VBF, $WH$, $ZH$ and $t\bar{t}H$ and people have obtained consistent results in the search of Higgs in those processes. In this work, we will analyze the production of Higgs boson in association with a
single top quark ($tH$) in proton-proton collisions with the CMS experiment of the LHC. The motivation of the work is this mechanism of production of the Higgs boson has not been observed before by any experiment. The exploration of Higgs production on the tH channel is subject relatively new. Many channels are yet out of reach experimentally. At the same time, many theoretical
proposals are yet to be put on test. It is possible that studies on the Higgs boson give evidence of small deviations from the SM  and test new physics that are beyond SM such as String Theory and Supersymmetry. \\
%********************************** % Third Section  *************************************

@@

For the leptons we have electrons ($e$), muons ($\mu$), and taus ($\tau$) with charge -1 and their neutrinos $\nu_e, \nu_\mu,\nu_\tau$, are chargeless particles. Leptons have their respective antiparticle; same structure with opposite charge. All of them have a spin (intrinsic angular momentum) of $\frac{1}{2}$\cite{griff}.\\

 Quarks combine to form composite particles called hadrons. Quarks cannot be found in isolation, they always exist in form of hadrons that can be mesons or baryons. 
Mesons , are bosons made up of a quark-antiquark pair, such as pion ($\pi$) and kaon ($K$). Mesons were theorized by Yukawa in 1937 and discovered in 1947 by Powell\cite{griff}.
Baryons are fermions made of 3 quarks. Examples of baryons are protons and neutrons. Both particles can interact via strong force.\\
Quarks were first theorized by Murray Gell-Mann and 
George Zweig in 1964, were introduced as parts of an ordering scheme for hadrons, and there was little evidence for their physical existence until scattering experiments at the Stanford Linear Accelerator Center in 1968\cite{griff}.
There are 6 of them: up (1968), down (1968), strange (1968), charm (theorized in 1970, discovered 1974), bottom (theorized 1973, discovered 1977), top (theorized 1973, discovered 1995).

Quarks have characteristics such as  electric charge, mass, color charge, and spin. The quarks up, charm and top have in common a charge of $\frac{2}{3}$, while the down, strange and bottom quarks have charge of $-\frac{1}{3}$. Among them, the top quark is the most massive of all, while the up and down are the least massive. 
The lepton and quarks have their respective antiparticle, which have the same mass, but inverted charge.\\  
Some of the particles in the SM are not stable, so they decays and form particles with lower masses. Other particles do not decay or their lifetimes are very long. The final states can be represented with a Feynman diagram according to the SM vertices\cite{mark}.

@@

The last group of the family are the bosons,which have a spin 1 and 0. 
There are 5 bosons: $W^{\pm}$, $Z$ with neutral charge are responsible for weak force; photon, a massless particle, is responsible for electromagnetic force; gluons $g$ which also massless, is related to strong force and Higgs boson, recently discovered , is in charge of give the property of mass to particles via Higgs mechanism. Bosons let other particles to interact, even themselves, which is shown in the figure \ref{sm}. 
\\
The $W$, $Z$ and Higgs bosons, can interact with the most of the particles, while the gluon can only interact with quarks and itself.
Those bosons act as mediators in decays and creation of complex
particles, as mesons and baryons.\\

@@

\begin{table}[!htbp]
\caption[Table of particles in SM]{Table of particles in SM\cite{pd}. Stable means no decay, or lifetime almost infinite. Particles speeds at 0.998c}
\renewcommand{\arraystretch}{1.5}
\begin{tabular}{|c|c|c|c|c|p{2.5cm}|}
\hline 
Particle	& Mass (MeV/$c^2$)  &Charge  & spin &Lifetime (s)  & Distance in lifetime (meters) \\ 
	\hline 
Up ($u$)	& 2.2 & $\frac{2}{3}$ & $\frac{1}{2}$ & stable &  -\\ 
	\hline 
Charm ($c$)	& 1280 &$\frac{2}{3}$  &  $\frac{1}{2}$ & $ 1.1 \times 10^{-12}$ &  5.21$\times 10^{-3}$ \\ 
	\hline 
Top	($t$)& 173100 & $\frac{2}{3}$ & $\frac{1}{2}$  & 5$\times$10$^{-25}$ &2.37$\times 10^{-15}$   \\ 
	\hline 
Down ($d$)	& 4.6 &$-\frac{1}{3}$  & $\frac{1}{2}$  & Stable & - \\ 
	\hline 
Strange ($s$)	& 96 &$-\frac{1}{3}$  & $\frac{1}{2}$  &$1.24 \times 10^{-8}$  & 58.7 \\ 
	\hline 
Bottom ($b$)	& 4180 &$-\frac{1}{3}$  &  $\frac{1}{2}$ &$1.3 \times 10^{-12}$   & 6.16 $\times 10^{-3}$\\ 
	\hline 
$W$ 	& 80379 &$\pm$1  & 1 & 3$\times$ 10$^{-25}$ &  1.42$\times 10^{-15}$\\ 
	\hline 
$Z$ & 91187.6 &0  & 1 & 3$\times$ 10$^{-25}$  &1.42$\times 10^{-15}$ \\ 
\hline
Photon ($\gamma$) & 0 &0  &  1&Stable  & - \\ 
\hline
Gluon ($g$)	& 0 &0  &  1&Stable  & - \\ 
	\hline 
Higgs ($H$)	& 125.18 &0  &  0& 1.56 $\times$ $10^{-22}$ & 7.39$\times 10^{-13}$ \\ 
	\hline 
Electron ($e$)& 0.511 & -1 &   $\frac{1}{2}$& Stable & - \\ 
	\hline 
Muon ($\mu$)	& 105.7 & -1  &  $\frac{1}{2}$ & 2.2$\times$10$^{-6}$ & 10419.85 \\ 
	\hline 
$\tau$	& 1776.86 &-1  &  $\frac{1}{2}$ & 2.9$\times$10$^{-13}$ & 1.37$\times 10^{-3}$\\ 
	\hline 
$\nu_e$	$\nu_\mu$ $\nu_\tau$& 0 & 0 &  $\frac{1}{2}$ & Stable &  -\\
	\hline 
\end{tabular} 
\label{SM table}
\end{table}
\newpage

\section{Higgs Boson}
The Higgs is a boson with spin 0 and chargeless. First theorized in 1964 by a group of theoretical physicist composed by Peter Higgs, Fran\c cois Englert, Robert Brout, Gerald Guralnik, C. Richard Hagen, and Tom Kibble in separated groups. They formulated the Higgs mechanism that explain the generation of mass for the particles. \\

@@

\section{Top quark}
The top quark  is the most massive particle  with a charge of 2/3 with spin of 1/2. First proposed by  M. Kobayashi and T. Maskawa in 1973 to explain observed CP violations in Kaon decay\cite{griff}. On 1995, in the Tevatron at Fermilab in Illinois, the top quark  was discovered via proton antiproton collisions with center of mass energy between 1.8 to 1.96 TeV in the CDF and D0 experiments.The top quark production process was via ggF, which generated a pair of $t\bar{t}$\cite{top}.
Since then,the properties of the top quark have been studied at the Tevatron Collider at Fermilab, and by the ATLAS and CMS experiments at the Large Hadron Collider (LHC) at CERN.\\
However, the top quark often receives special attention in new physics models because its mass requires near unity coupling to the Higgs boson\cite{top}.
As shown in table \ref{SM table}, the top quark lifetime is so small that they decay before they can hadronise, that is to say, the process to create hadrons from quarks and gluons.\\

@@

$W^i$ (i=1,2,3) are three  SU(2) gauge bosons and B is a U(1) gauge boson\footnote{$\epsilon^{ijk}$ represents the Levi-Civita tensor, where $\epsilon^{ijk}= \begin{dcases}
	+1 \quad  \text{for even permutation} \\
	-1 \quad \text{for odd permutation}\\
	0 \quad \text{i=j, or j=k, or k=i } 
	\end{dcases}$.}. In the end, those gauge bosons, after introduce the Higgs mechanism  become the $W^\pm$ , $Z$ and $\gamma$ vector boson
\cite{ew1}\cite{ew2}. %Z bosons  and photons will be formed by a combination of W and B fields.\\
\\

@@

We define the physical vector boson $W^-_\mu$. $W^+_\mu$ and $Z$ 
\begin{align} 
W^{\pm}_\mu=\frac{1}{\sqrt{2}} (W^1_\mu \mp iW^2_\mu) \label{smo}\\ 
Z_\mu=\frac{1}{\sqrt{g^2+g'^2}}\left(gW^3_\mu -g'B_\mu   \right) \label{zboson}
\end{align}

@@

The last part of the lagrangian is the Yukawa lagrangian
\begin{align}
\mathcal{L}_\text{yukawa}=\sum_{m,n}^{3}  \Gamma^u_{mn}\bar{q}_{m\text{,}L} \tilde{\Phi} u_{n\text{,}R}+\Gamma^d_{mn}\bar{q}_{m\text{,}L} \Phi d_{n\text{,}R}+\Gamma^e_{mn}\bar{l}_{m\text{,}L} \Phi e_{n\text{,}R}+h.c
\end{align}
The matrices $\Gamma_{mn}$ describe the so called Yukawa couplings between Higgs doublet $\Phi$ and the fermions.\footnote{Here $
	\tilde{\Phi}=i\sigma^2 \Phi^\dagger =\left(\begin{array}{c}
	\phi^{0^\dagger} \\
	-\phi^-
	\end{array} \right)$
	% \Phi =\sqrt{-\frac{\mu^2}{2\lambda}}=\frac{v}{\sqrt{2}}  \qquad \lambda >0
	 is necessary for the correct transformation in the case of up quarks. \\
h.c refers to hermitian conjugate terms.  }  The indices  m and n mean sum over the families. \\
%It is required to have two representations of Higgs fields
%with $Y=\frac{1}{2}$  and $\frac{1}{2}$ to give masses to  quarks and electrons.

@@

In tH process, Higgs boson is radiated from a single top quark as shown in \ref{psu}. $tH$ process is a new field of study, given that there are few experiments and its production rate (probability) is small, so the detection of Higgs bosons in this channel opens to new discoveries.
\\

@@

In the table \ref{crt} shows the Higgs production cross section for pp collision and the number of events produced using a integrated luminosity of 35.9 fb$^{-1}$, that is the luminosity measured for the Run 2 from LHC on 2016\cite{pd}.

@@

Here we can see ggF process have the biggest cross section of all Higgs mechanism and generates the greatest number of events possibles. For the processes $tH$ and $t\bar{t}H$, they have the smallest cross section, so the probability for the processes is low and generates a small number of events. 
\pagebreak
\footnotetext{Taken from \url{ https://twiki.cern.ch/twiki/bin/view/LHCPhysics/LHCHXSWGCrossSectionsFigures}}
\section{Higgs decay rates}
In particle physics,two important properties are the lifetime and the decay rates of a particle. The lifetime is related to the total decay rate $\Gamma$, that is the probability per unit of time a particle decay, 
\begin{align}
    dN=-\Gamma N dt
\end{align}
With N the original number of particles before the decay. From here is evident that number of particles left in a decay is 
\begin{align}
N(t)=N_0 e^{-\Gamma t}
\end{align}
